---
title: "Quantitative Content Analysis: Lecture 2"
author: "Olga Kononykhina"
date: "22 Februar 2017"
output:
  beamer_presentation:
    theme: "CambridgeUS"
    colortheme: "default"
    fonttheme: "default"
    fig_caption: false
  ioslides_presentation:
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Today's Outline

**Intro to R (Part II)**

- Vectors, matrices, data frames, lists
- Working with data sets
- Loops and conditions
- Creating and applying functions

## Vectors

A vector is a container of objects put together in an order.
```{r, echo = T}
# Define a vector
a <- c(1,4,5)
b <- c(3,6,7)
```

```{r, echo = T}
# Join multiple vectors
ab <- c(a,b)
```

```{r, echo = T, eval = F}
# Find vector length (number of its elements)
length(a)
```

```{r, echo = T, eval = F}
# Reference vector components
ab[4] # Index one element in vector
ab[4:6] # Index several elements in a vector
ab[ab==6] # Index with condition
```


## Operations on vectors

| Operation     | Meaning                  |
| --------      | ------------------------ |
| `sort(x)`     | sort a vector            |
| `sum(x)`      | sum of vector elements   |
| `mean(x)`     | arithmetic mean          |
| `median(x)`   | median value             |
| `var(x)`      | variance                 |
| `sd(x)`       | standard deviation       |
| `factorial(x)`| factorial of a number    |


## Task 1

Calculate the mean of the vector 1,99,3,4,5,6,8. What‘s the mean if you out the 'outlier'?


## Task 1 (solution)

Calculate the mean of the vector 1,99,3,4,5,6,8. What‘s the mean if you out the 'outlier'?

```{r, echo = T}
# Defining vector using sequence between 3 and 6
a <- c(1,99,3:6,8) 
mean(a)

# Calculate the mean of a but exclude the highest number
mean(a[a!=max(a)])
```


## Matrices

A Matrix is a square 2 dimensional container, i.e. vectors combined by row or column

- Must specify number or rows and columns `matrix(x,nrow,ncol,byrow)`

    + x: vector of length nrow*ncol
    + nrow: number of rows
    + ncol: number of columns
    + byrow: TRUE or FALSE, specifies direction of input


## Task 2

Assign a 6 x 10 matrix with 1,2,3,…,60 as the data.


## Task 2 (solution)

Assign a 6 x 10 matrix with 1,2,3,…,60 as the data.

```{r, echo = T}
m <- matrix(1:60, nrow=6, ncol=10)
m
```


## Referencing matrices

- Like vectors, you can reference matrices by elements

    + `a[i,j]` refers to the ith row, jth column element of object `a`

- Can also reference rows/columns, these are vectors

    + `a[i,]` is ith row, `a[,j]` is jth column


## Task 3

Extract the 9th column of the matrix from the previous problem. How can you find the 4th element in the 9th column?
    
    
## Task 3 (solution)

Extract the 9th column of the matrix from the previous problem. How can you find the 4th element in the 9th column?

```{r, echo = T}
m[,9]
m[4,9]
m[,9][4]
```


## Data frames

Data frames are a two-dimensional container of vectors with the same length. Each column (vector) can be of a different class and can be indexed with `[,]` or `$`. You can use functions like `nrow()`, `ncol()`, `dim()`, `colnames()`, or `rownames()` get information about your df.

```{r, echo = T}
# Combine two vectors into a data frame
number <- c(1, 2, 3, 4)
name <- c('John', 'Paul', 'George', 'Ringo')
df <- data.frame(number, name, stringsAsFactors = FALSE)
df
```


## Lists

A list is an object containing other objects that can have different lengths and classes.

```{r, echo = T}
# Create a list with three objects of different lengths
list1 <- list(beatles = c('John', 'Paul', 'George', 'Ringo'),
              alive = c('Paul', 'Ringo'), albums = 1:13)
list1
```


## R's build-in data sets
There are a number of example data sets available within `R`.

```{r, echo = T, eval=FALSE}
# List internal data sets:
data()
```

```{r, echo = T, eval = F}
# Load swiss data set:
data(swiss)
# Find data description:
?swiss
```


## Importing data

You can read  data and assign it to an object. The most frequently used functions to read data include:

- `load()`: To open `.RData` files
- `read.csv()`: Reads csv file
- `read.table()`: Is more general and allows to set separators
- `read.dta()`: From `foreign` library, used to read Stata files


## Exporting data

You can export your data in various formats:

- `save()`: Only readable in R, but can store multiple objects
- `write.csv()`: Writes matrix/dataframe to csv
- `write.table`: Writes matrix to a tab delimited text file
- `write.dta()`: Writing in Stata format, requires `foreign` library


## For() loops

For() loops are used to loop around a vector/matrix to do something.
```{r, echo = T}
m <- matrix(1:5, nrow=1, ncol=5)
m
for (j in 1:3){
     m[,j]=0
}
m
```


## For() loops (II)

You can also 'nest' a `for()` loop in another `for()` loop
```{r, echo = T}
m <- matrix(1:15, nrow=3, ncol=5)
for (i in 1:2){
  for (j in 1:4){
    m[i,j]=0
  }
}
m
```


## Task 4

Create a matrix with `matrix(1:20,nrow=4,ncol=5)` and another with `matrix(NA,nrow=4,ncol=5)`. Adapt the second to the first matrix using `for()`

- hint: define a 'counter‘ variable that increments by 1 each time you move to the next cell


## Task 4 (solution)

Create a matrix with `matrix(1:20,nrow=4,ncol=5)` and another with `matrix(NA,nrow=4,ncol=5)`. Adapt the second to the first matrix using `for()`


```{r, echo = T}
counter=1
m1 <- matrix(1:20,nrow=4,ncol=5)
m2 <- matrix(NA,nrow=4,ncol=5)
for (j in 1:5){
  for (i in 1:4){
    m2[i,j]=counter
    counter=counter+1
  }
}
```

## Task 4 (alternative solution)

Create a matrix with `matrix(1:20,nrow=4,ncol=5)` and another with `matrix(NA,nrow=4,ncol=5)`. Adapt the second to the first matrix using `for()`


```{r, echo = T}
m1 <- matrix(1:20,nrow=4,ncol=5)
m2 <- matrix(NA,nrow=4,ncol=5)
for (j in 1:5){
  for (i in 1:4){
    m2[i,j]=m1[i,j]
  }
}
```


## If() statements 

If() statements are used to make conditions on executing some code. If condition is true, action is done. 
```{r, echo = T}
a <- 3
b <- 4
number <- 0
if(a<b){
  number=number+1
}
number
```

Tests for conditions: `==; >; <; >=; <=; !=`


## Task 5

Create the two objects `a <- sample(c(4,0),20,replace=TRUE)` and 
`m <- matrix(a,nrow=4,ncol=5)`. Now recode all the 4s into 1s using if() and for() statements.


## Task 5 (solution)

Create the objects `a <- sample(c(4,0),20,replace=TRUE)` and 
`b <- matrix(a,nrow=4,ncol=5)`. Now recode all the 4s into 1s in `b` using if() and for() statements.

```{r, echo = T}
a <- sample(c(4,0),20,replace=TRUE)
b <- matrix(a,nrow=4,ncol=5)

for (j in 1:5){
  for (i in 1:4){
    if (b[i,j]==4){
      b[i,j]=1       
    }
  }
}
```


## Creating Functions

If you want to repeat an operation it is often useful to create your own Function. Functions have names, inputs and outputs and simplify your code.

```{r, echo = T}
# Find the sample mean of a vector
fun_mean <- function(x){
    sum(x) / length(x)
}

## Find the mean
data(swiss)
fun_mean(x = swiss$Infant.Mortality)
```


## Task 6

Write a function that takes a number and doubles it.


## Task 6 (solution)

Write a function that takes a number and doubles it.

```{r, echo = T}
double <- function(x){
output <- x * 2
output
}
double(8)
```


## Apply function

`Apply` allows you to apply a function to every row or every column. This can be done with a `for()` loop, but `apply()` is usually much faster and simpler. `Apply()` takes the following form: `apply(X, MARGIN, FUN, ...)`.

```{r, echo = T}
m <- matrix(c(1:10, 11:20), nrow = 10, ncol = 2)
# mean of the rows
apply(m, 1, mean)
# mean of the columns
apply(m, 2, mean)
```


## Task 7

Load up the build-in R dataset 'iris' and use `apply()` to get the mean of the first 4 variables, by species.


## Task 7 (solution)

Load up the build-in R dataset 'iris' and use `apply()` to get the mean of the first 4 variables.

```{r, echo = T}
attach(iris)
apply(iris[,1:4], 2, mean)
```


## Packages

You can greatly expand the number of functions by installing and loading 
user-created packages.

```{r, eval=FALSE}
# Install dplyr package
install.packages('dplyr')

# Load dplyr package
library(dplyr)
```

You can also call a function directly from a specific package with the double
colon operator (`::`).

```{r, eval=FALSE}
Grouped <- dplyr::group_by(combined_df, character_vector)
```


## Piping

Piping allows to pass a value forward to a function call and produces faster compilation and enhanced code readability. In `R` use `%>%` from the `dplyr` package.


```{r, echo = T}
# Not piped:
values <- rnorm(1000, mean = 10)
value_mean <- mean(values)
round(value_mean, digits = 2)
```

```{r, echo = T, message=F}
# Piped:
library(dplyr)
rnorm(1000, mean = 10) %>% mean() %>% round(digits = 2)
```

## Help

- Type help() (`or ?`) to see documentation 
- Read Wickham & Grolemund's _R For Data Science_ Handbook
- Check out the help function for a couple of functions used in today's course to see 'what is what' in the documentation

## Next week

- Tidy data
- Data manipulation with `dplyr`
- Visualizations with `ggplot`